---
title: "p8105_hw5_kr3145"
author: "Kallan Roan"
date: "2025-11-08"
output: github_document
---

Load libraries 

```{r}
library(tidyverse)
```

Preset settings

```{r}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

```

set seed

```{r}
set.seed(1)
```


## Problem 1

Create function

```{r}
bday_sim = function(n_room) {
  
  birthdays = sample(1:365, n_room, replace = TRUE)
  
  repeated_bday = length(unique(birthdays)) < n_room
  
  repeated_bday
  
}
```

Create simulation for group size between 2 and 50

```{r}
sim_results_df = 
  expand_grid(
    sample_size = 2:50, 
    iter = 1:10000
  ) |> 
  mutate(
    results = map(sample_size, bday_sim)
  ) |> 
  unnest(results)
```

Use simulation results to calculate probability and plot

```{r}
sim_results_df |> 
  group_by(sample_size) |> 
  summarize(
    prob = mean(results)
  ) |> 
  ggplot(aes(x = sample_size, y = prob)) +
  geom_point() +
  geom_smooth()
```

The plot shows that as the group size increases, the probability that at last 2 people in the group will share a birthday will also increase. The increase starts to plateau slightly as it gets to group size of 50. 

## Problem 2



